---
title: "Individual1"
author: "Caroly Coronado- Vargas"
date: "10/25/2021"
output: html_document
---
## Part one: Cleaning Data ##

```{r}
library(dplyr)
library(tidyverse)
library(readxl)
defense <- read_excel('../Lab4/cyclonesFootball2020.xlsx', sheet='Defensive')

offensive <- read_excel('../Lab4/cyclonesFootball2020.xlsx', sheet='Offensive')

biography <- read_excel('../Lab4/cyclonesFootball2020.xlsx', sheet='Biography')

str(defense)
str(offensive)
str(biography)
```

### 1 ###

```{r}
defClean <- defense %>% 
  rename(Opponent = Opponent_Opponent) %>% 
  mutate(Opponent = factor(Opponent,                       as.character(unique(Opponent)))) %>% 
  
  rename(Player = Name) %>% 
  mutate(Player = factor(Player,                       as.character(unique(Player))))

str(defClean)

offClean <- offensive %>% 
  rename(Opponent = Opponent_Opponent) %>% 
  mutate(Opponent = factor(Opponent,                       as.character(unique(Opponent)))) %>% 
  
  rename(Player = Name) %>% 
  mutate(Player = factor(Player,                       as.character(unique(Player))))

str(offClean)
```

### 2 ###

```{r}
offClean <- offClean %>%
  dplyr::mutate(across(Receiving_REC:Passing_INT, as.numeric))

str(offClean)

defClean <- defClean %>%
  dplyr::mutate(across(Tackles_Solo:Pass_PB, as.numeric))

str(defClean)
```
## 3 ##

```{r}
str(biography)

bioClean<- biography %>%
  separate(Height,c('Height.ft', 'inches'), sep = '-', convert = TRUE, remove = TRUE) %>%
  mutate(Height.ft = 12*Height.ft + inches) %>%
  select(-inches)

str(bioClean)
```
### 4 ###

```{r}
defClean<- defClean%>%
  group_by(Player,Opponent)%>%
  mutate(gameNum= row_number())
  
defClean

offClean<- offClean%>%
  group_by(Player,Opponent)%>%
  mutate(gameNum= row_number()) 

  
offClean

str(defClean)
str(offClean)
str(bioClean)

```

## Part Two: Tidying and Exploring ##

### 1 ###

```{r}
library(reshape2)
offClean

offClean<- pivot_longer(offClean,Receiving_REC:Receiving_TD, names_to = 'StatType',values_to = 'Stat')

```
### 2 ###

```{r}
offClean
g<- offClean%>%
  group_by(StatType,Player)%>%
  summarize(SumStat=sum(Stat))

g

```
### 3 ###

*We can see that most players have higher Receiving Yards followed by Recieving Touch Downs.*

```{r}
library(ggplot2)

bp<- ggplot(g, aes(x=StatType, y=SumStat), group=StatType)+geom_boxplot(aes(fill=StatType))+ facet_wrap(~Player)

bp
```

### 4 ###

*ISU had better offense against Oklahoma (first game) than Oregon since they had larger number of receiving yards.*

```{r}
offClean
Opps<-c("Oregon","Oklahoma")

offClean1<-offClean %>% 
  filter(Opponent %in% Opps)%>%
  filter(gameNum==1)%>%
  pivot_wider(names_from =StatType ,values_from =Stat)%>%
  group_by(Opponent)%>%
  summarize(Rec_YDS=sum(Receiving_YDS,na.rm=TRUE))
  
offClean1

ggplot(offClean1, aes(x=Rec_YDS, y=Opponent))+ geom_point()
```

### 5 ###

```{r}
bioClean

bioClean<- bioClean%>%
  separate(Hometown, c('City', 'State'), sep=',')
bioClean
```

### 6 ###

```{r}
NumState<-bioClean%>%
  group_by(State)%>%
  summarise(count_State=n())
NumState
```

### 7 ###

*Based on our graphs, When QB Brock Purdy has higher Passing Yards, the team has higher Rushing Touch Downs, thus contributing to the teams overall performance in a positive way, Looking at the defense does not make sense in our data since Purdy plays offense. looking at our graph we can see no correltaion*

```{r}
#Offense

BrockOff<-offClean%>%
  filter(Player=="Purdy, Brock")%>%
  group_by(Opponent)%>%
  summarize(Pass_Yards=sum(Passing_YDS))
BrockOff

TeamOff<-offClean%>%
  group_by(Opponent)%>%
  summarize(Rushing_Td=sum(Rushing_TD,na.rm = TRUE))
TeamOff

df<-full_join(BrockOff, TeamOff)
df

ggplot(df, aes(x=Pass_Yards, y=Rushing_Td))+ geom_point()


#Defense
defClean
BrockDef<-defClean%>%
  filter(Player=="Purdy, Brock")%>%
  group_by(Opponent)%>%
  summarize(Tackles_solo=sum(Tackles_Solo,na.rm = TRUE))
BrockDef

TeamDef<-defClean%>%
  group_by(Opponent)%>%
  summarize(tackles_TFL=sum(Tackles_TFL,na.rm = TRUE))
TeamDef

df2<-full_join(BrockDef, TeamDef)
df2

ggplot(df2, aes(x=Tackles_solo, y=tackles_TFL))+ geom_point()
```

